---
title       : "EMTF $p_T$ Regression"
subtitle    :
author      : Khristian Kotov
job         :
framework   : io2012        # {io2012, html5slides, shower, dzslides, ...}
highlighter : highlight.js  # {highlight.js, prettify, highlight}
hitheme     : tomorrow      #
github:
  user: kkotov
  repo: emtfPtRegression
url:
  lib:    ../../../../monohiggs/plots/libraries
widgets     : [mathjax]     # {mathjax, quiz, bootstrap}
mode        : selfcontained # {standalone, draft}
---

## Outline

1. New input data

2. Model selection
  - 

3. Trees
  - boosting (BDT), bagging (Random Forest)
  - predictors' importance
  - residuals and turn-on curves
  - ROCs and efficiency vs. rate curves

--- &twocol

## New Monte Carlo simulation data 

* Andrew's <span style="font-weight:bold;color:brown">/store/user/abrinke1/EMTF/MC/SingleMu_Pt1To1000_FlatRandomOneOverPt</span>
  - a double muon sample, flat in $0 < |\eta| < 2.5$ (barrel included)
  - about 300$k$ muons in endcaps extending to $p_T = 1000~GeV$, flat in $log(p_T)$

```{r message = F, echo = F}
mode_inv <- 15
mode = c(0, 0, 12, 0, 10, 5, 14, 0, 9, 5, 13, 3, 11, 7, 15)
df <- read.csv(file="../../SingleMu_Pt1To1000_FlatRandomOneOverPt.csv",header=T,sep=',')

d1 <- df[df[,"mode.0."]==mode[mode_inv], c( grep("\\.[0-1]\\.",colnames(df),invert=T) , grep(".0.",colnames(df),fixed=T) ) ]
d2 <- df[df[,"mode.1."]==mode[mode_inv], c( grep("\\.[0-1]\\.",colnames(df),invert=T) , grep(".1.",colnames(df),fixed=T) ) ]

colnames(d1) <- sub(".0.", "", colnames(d1),fixed=T)
colnames(d2) <- sub(".1.", "", colnames(d2),fixed=T)

d <- rbind(d1,d2)

require(ggplot2)
```

*** =left

```{r, fig.height=5., fig.width=5., fig.align = 'center', comment = NA, message = F, echo = F}
ggplot(df, aes(x=log(muPtGen) )) +
    geom_histogram(breaks=seq(0,7,0.1)) +
    theme(
        title = element_text(size=20),
        axis.title.x = element_text(size=20),
        axis.text.x  = element_text(size=15)
    ) +
    labs( x=expression(log(p[T])),
          y="count",
          title=expression(paste("Transformed ", p[T], " spectrum"))
    )
```

*** =right

```{r, fig.height=5., fig.width=5., fig.align = 'center', comment = NA, message = F, echo = F}
ggplot(df, aes(x=muEtaGen)) +
    geom_histogram(breaks=seq(-2.6,2.6,0.1)) +
    theme(
        title = element_text(size=20),
        axis.title.x = element_text(size=20),
        axis.text.x  = element_text(size=15)
    ) +
    labs( x=expression(eta),
          y="count",
          title="Pseudorapidity spectrum"
    )
```

*** =fullwidth

--- &twocol

## Model with all predictors included


```{r, comment = NA, message = F, echo = F}

require(ranger)
vars <- with(d,data.frame( 1/muPtGen,
                           2*atan(exp(-muEtaGen)),
                           pt,
                           mypt,
                           dPhi12,
                           dPhi13,
                           dPhi14,
                           dPhi23,
                           dPhi24,
                           dPhi34,
                           dTheta12,
                           dTheta13,
                           dTheta14,
                           dTheta23,
                           dTheta24,
                           dTheta34,
                           factor(clct1,levels=c(2,3,4,5,6,7,8,9,10)),
                           factor(clct2,levels=c(2,3,4,5,6,7,8,9,10)),
                           factor(clct3,levels=c(2,3,4,5,6,7,8,9,10)),
                           factor(clct4,levels=c(2,3,4,5,6,7,8,9,10)),
                           factor(fr1,levels=c(0,1)),
                           factor(fr2,levels=c(0,1)),
                           factor(fr3,levels=c(0,1)),
                           factor(fr4,levels=c(0,1))
                           )
            )
predictors <- c("dPhi12", "dPhi13", "dPhi14", "dPhi23", "dPhi24", "dPhi34", "dTheta12", "dTheta13", "dTheta14", "dTheta23", "dTheta24", "dTheta34", "clct1", "clct2", "clct3", "clct4", "fr1", "fr2", "fr3", "fr4")
colnames(vars) <- c("muPtGenInv", "theta", "ptTrg", "mypt", predictors )
predictors <- c("theta", predictors)

set.seed(1)

part <- sample(seq(nrow(vars)), as.integer(nrow(vars)*0.75), replace=F)
trainSet <- vars[part,]
testSet <- vars[-part,]
POI <- which(colnames(vars)=="muPtGenInv")

f <- as.formula(paste("muPtGenInv ~ ", paste(predictors, collapse= "+")))

invisible( capture.output( modelFit <- ranger(f, data=trainSet, importance="impurity") ) )
```

*** =left

```{r, fig.height=5., fig.width=5., fig.align = 'center', comment = NA, message = F, echo = F}
source("../turnOn.R")
turnOn(modelFit,shift=14)
```

*** =right

```{r, fig.height=5., fig.width=5., fig.align = 'center', comment = NA, message = F, echo = F}
barplot( modelFit$variable.importance, horiz=T, col='red', xlab='Delta MSE', las=2 )
```

*** =fullwidth

--- .class #id

## qwe
