---
title       : "EMTF $p_T$ Regression"
subtitle    :
author      : Khristian Kotov
job         :
framework   : io2012        # {io2012, html5slides, shower, dzslides, ...}
highlighter : highlight.js  # {highlight.js, prettify, highlight}
hitheme     : tomorrow      #
github:
  user: kkotov
  repo: emtfPtRegression
widgets     : [mathjax]     # {mathjax, quiz, bootstrap}
mode        : selfcontained # {standalone, draft}
---

## Outline (continue of [previous talk](https://kkotov.github.io/emtfPtRegression/2017.02.10/))

Measured rate spectrum
* fitting with a power spectrum
* different components

Low $p_T$ rate spectrum component
* closer look
* better parametrization

Deconvolution of the "true" underlying rate from the spectrum
* constructing the likelihood
* closure tests


--- &twocol

## Properties of the measured rate

```{r message = F, echo = F}
mode_inv <- 15
mode = c(0, 0, 12, 0, 10, 5, 14, 0, 9, 5, 13, 3, 11, 7, 15)
df <- read.csv(file="../../pt/SingleMu_Pt1To1000_FlatRandomOneOverPt.csv",header=T,sep=',')
d1 <- df[df[,"mode.0."]==mode[mode_inv], c( grep("\\.[0-1]\\.",colnames(df),invert=T) , grep(".0.",colnames(df),fixed=T) ) ]
d2 <- df[df[,"mode.1."]==mode[mode_inv], c( grep("\\.[0-1]\\.",colnames(df),invert=T) , grep(".1.",colnames(df),fixed=T) ) ]
colnames(d1) <- sub(".0.", "", colnames(d1),fixed=T)
colnames(d2) <- sub(".1.", "", colnames(d2),fixed=T)
d <- rbind(d1,d2)

source("../metrics.R")
binning <- seq(0,100,2)
nBins   <- length(binning)
slimDF  <- with(d, data.frame(true_pT = muPtGen, response_pT = pt) )
slimDF$trueBin <- sapply(slimDF$true_pT, findBin, binning)
pSpec   <- aggregate(slimDF, by=list(bin=slimDF$trueBin), function(x) x )
norms   <- sapply(pSpec$response_pT,length)
counts  <- t(sapply(pSpec$response_pT,function(x) sapply(binning,function(y) sum(unlist(x)>y) )))
turnOn  <- counts/norms

for(f in paste("../../pt/s",c(0,1,2,3),".RData",sep='')) load(f)
zb <- s0
zb <- rbind(zb,s1)
zb <- rbind(zb,s2)
zb <- rbind(zb,s3)

ptMes <- zb$ptEMTF

#dfXY <- data.frame( x=runif(1e8,1,1000), y=runif(1e8) )
#pt <- dfXY[ 1/dfXY$x^2 > dfXY$y,"x"]
#rm(dfXY)
#qqplot(runif(30000,0.001,1)^-1,pt) # for the power spectrum this is much more efficient generator
## also check this out:
# pt <- dfXY[ 1/dfXY$x^4 > dfXY$y,"x"]
# qqplot(runif(30000,0.00001,1)^(-1/3),pt)
# so the scale-formula is b = 1/(a+1), e.g. a = -4 => b = -1/3, a = -2 => b = -1

pt2.7 <- runif(30000000,0.00001,1)^(-1/1.7)
rate2.7 <- sapply(binning[2:nBins], function(x) sum(pt2.7 > x) )

require(ggplot2)
```

*** =left

```{r, fig.height=5., fig.width=5., fig.align = 'center', comment = NA, message = F, echo = F}

ratePlot <- ggplot() +
        geom_freqpoly( data=data.frame(pT=pt2.7),
                        breaks=seq(2,257,2),
                        position = "identity",
                        aes( x = pT,
                             y = (..count..)/sum(..count..),
                             colour = "r"
                        )
                      ) +
        geom_freqpoly( data=data.frame(pT=ptMes),
                        breaks=seq(2,257,2),
                        position = "identity",
                        aes( x = pT,
                             y = (..count..)/sum(..count..),
                             colour = "b"
                        )
                      ) +
        theme(
            title = element_text(size=20),
            axis.title.x = element_text(size=15),
            axis.text.x  = element_text(size=20),
            legend.position = c(.6, .7),
            legend.background = element_rect(fill = 'grey92', colour = 'black', size=0),
            legend.text=element_text(size=rel(1.5)),
            legend.title=element_text(size=rel(0.8), face="bold", hjust=0)
        ) +
        labs( x=expression(paste(p[T], " (GeV/c)")),
              y="density",
              title="Rate"
        ) +
        scale_colour_manual(name = "", values=c("r" = "red", "b"="blue"), labels=c("b"="EMTF", "r"=expression(1/p[T] ^2.7))) +
#        scale_fill_manual(name = "",   values=c("r" = "red", "b"="blue"), labels=c("b"="EMTF", "r"=expression(1/p[T] ^2.7))) +
        scale_y_log10()

invisible( capture.output( suppressWarnings( ratePlot ) ) )
```

*** =right

```{r, fig.height=5., fig.width=5., fig.align = 'center', comment = NA, message = F, echo = F}
#qqplot(pt0.45,ptMes)
# ggplot2 equivalent of the build-in above:

qqPlot <- ggplot( data.frame(qqplot(x=ptMes, y=pt2.7, plot.it=FALSE)) ) +
        geom_point(aes(x=x, y=y)) +
        theme(
            title = element_text(size=20),
            axis.title.x = element_text(size=20),
            axis.text.x  = element_text(size=20),
            axis.title.y = element_text(size=20),
            axis.text.y  = element_text(size=0)
        ) +
        labs( y=expression(paste(1/p[T] ^2.7, " spectrum")),
              x=expression(paste("measured ", p[T], " rate (Gev/c)")),
              title="Distribution similarity (q-q)"
        ) #+ scale_x_log10()

qqPlot
```

*** =fullwidth

Clearly seen are two components meeting at $\sim$100 GeV/$c$ (and the EMTF pT saturation)

Since all of the thresholds of interest are < 100 GeV/$c$ let's zoom in to the first component

--- &twocol

## Low $p_T$ part of the measured rate 

```{r message = F, echo = F}
# let's fit a binned likelihood to get the answer quick

binning <- seq(2,80,1)
rateThr <- sapply(binning, function(x) sum(ptMes > x) )
rateBin <- sapply(1:(length(binning)-1), function(x) sum(ptMes > binning[x] & ptMes < binning[x+1]) )

binLikelihood <- function(par, rate, mode) {
    norm1     <- par[1]
    powerLaw1 <- par[2]
    norm2     <- par[3]
    powerLaw2 <- par[4]
    shift     <- par[5]
    if(mode == "aboveThr")
        estRate   <- sapply(binning,function(x) rate[1]*(norm1/exp(x*powerLaw1) + norm2/((x+shift)^powerLaw2)))
    else
        estRate   <- sapply(binning[1:(length(binning)-1)],function(x) rate[1]*(norm1/exp(x*powerLaw1) + norm2/((x+shift)^powerLaw2)))
#    sum( (rate - estRate)^2 ) # simple MSE for log likelihood, not the best choise, of course
#    sum( (rate - estRate)^2 / estRate ) # gaussian probability for likelihood works if all estRate >~5
    sum( estRate + rate*(log(rate/estRate) - 1) ) # poisson likelihood works for any estRate (!=0, of course)
}

#invisible( capture.output( ) )
suppressWarnings( fitAbove <- nlm(binLikelihood, c(1,2,1,1,10), rate = rateThr, mode="aboveThr", hessian=T) )
suppressWarnings( fitDiff  <- nlm(binLikelihood, c(1.8,1.2,17,3.1,3), rate = rateBin, mode="diff", hessian=T) )
#solve(fitMes$hessian) # to get covariance matrix if needed

norm1_d     <- fitDiff$estimate[1]
powerLaw1_d <- fitDiff$estimate[2]
norm2_d     <- fitDiff$estimate[3]
powerLaw2_d <- fitDiff$estimate[4]
shift_d     <- fitDiff$estimate[5]
approxRateDiff  <- sapply(binning[1:(length(binning)-1)],function(x) rateBin[1]*(norm1_d/exp(x*powerLaw1_d) + norm2_d/((x+shift_d)^powerLaw2_d)))

norm1_a     <- fitAbove$estimate[1]
powerLaw1_a <- fitAbove$estimate[2]
norm2_a     <- fitAbove$estimate[3]
powerLaw2_a <- fitAbove$estimate[4]
shift_a     <- fitAbove$estimate[5]
approxRateAbove <- sapply(binning,function(x) rateThr[1]*(norm1_a/exp(x*powerLaw1_a) + norm2_a/((x+shift_a)^powerLaw2_a)))

rateOrig_d <- data.frame(pT = binning[1:(length(binning)-1)], cnt = rateBin) 
rateFit_d  <- data.frame(pT = binning[1:(length(binning)-1)], cnt = approxRateDiff)

rateOrig_a <- data.frame(pT = binning, cnt = rateThr) 
rateFit_a  <- data.frame(pT = binning, cnt = approxRateAbove)
```

```{r message = F, echo = F}
makeRRplot <- function(rateOrig, rateFit, norm1, powerLaw1, norm2, powerLaw2, shift, title) {
        ggplot() +
        geom_line( data = rateOrig,
                   aes( x = pT,
                        y = cnt,
                        colour = "r"
                      )
                 ) +
        geom_line( data = rateFit,
                   aes( x = pT,
                        y = cnt,
                        colour = "b"
                   )
                 ) +
        labs( x=expression(paste(p[T], " (GeV/c)")),
              y="conuts",
              title=title
        ) +
        theme(
            title = element_text(size=20),
            axis.title.x = element_text(size=20),
            axis.text.x  = element_text(size=20),
            legend.position = c(.50, .85),
            legend.background = element_rect(fill = 'grey92', colour = 'black', size=0),
            legend.text=element_text(size=rel(1.5)),
            legend.title=element_text(size=rel(0.8), face="bold", hjust=0)
        ) +
        scale_colour_manual(name = "",
                            values = c("r" = "red", "b"="blue"),
                            labels = c("r"="EMTF",
                                       "b"=bquote(.(signif(norm1,2)) ~ "/" ~ e^ ~ {.(signif(powerLaw1,2)) ~ p[T]} ~ "+" ~ .(signif(norm2,2)) ~ "/(" ~ p[T] ~ "+" ~ .(signif(shift,2)) ~ ")"^.(signif(powerLaw2,2)) )
                                      )
                           ) +
        scale_y_log10()
}
```

*** =left

Simple power law is a crude approximation, exponent + power law works better:

```{r, fig.height=5., fig.width=5., fig.align = 'center', comment = NA, message = F, echo = F}
#makeRRplot(rateOrig_a, rateFit_a, norm1_a, powerLaw1_a, norm2_a, powerLaw2_a, shift_a, "Integrated rate above threshold")
makeRRplot(rateOrig_d, rateFit_d, norm1_d, powerLaw1_d, norm2_d, powerLaw2_d, shift_d, "Rate")
```

*** =right

But quantile-quantile plot is not straight $\rightarrow$ the fit is still rather mediocre, of course:

```{r, fig.height=5., fig.width=5., fig.align = 'center', comment = NA, message = F, echo = F}

ratePDF <- function(x, norm1, powerLaw1, norm2, powerLaw2, shift) norm1/exp(x*powerLaw1) + norm2/((x+shift)^powerLaw2)

pTvals <- seq(2,80,0.01)
rProbs <- ratePDF(pTvals, norm1_d, powerLaw1_d, norm2_d, powerLaw2_d, shift_d)

rateCDF <- cumsum(rProbs)

# invert CDF with approxfun to get quantile function
rateQuantiles <- approxfun(x=rateCDF/sum(rProbs), y=pTvals, yleft=2, yright=80)

# source https://github.com/tidyverse/ggplot2/blob/master/R/stat-qq.r turned out to be crucial for implementing qqPlot below

qqPlot <- ggplot( data.frame(x=sort(ptMes[ptMes<80 & ptMes>=2])), aes(sample=x)) +
        stat_qq(distribution = function(p) rateQuantiles(v=p) ) +
        labs( x="fit",
              y="measured",
              title="Quantile-quantile plot"
        ) +
        theme(
            title = element_text(size=20),
            axis.title.x = element_text(size=20),
            axis.text.x  = element_text(size=20)
        )

qqPlot
```

*** =fullwidth

qwe I'll stick with this parametrization keeping in mind the excluded [1-2] GeV/$c$ bin

--- .class #id

## Deconvolution of underlying rate spectrum

Construct likelihood $L(R|r,e) \sim -\prod_i Poiss(r_i | e_i ^j R_j)$ where, as [before](https://kkotov.github.io/emtfPtRegression/2017.02.10/#8),
* $r_i$ - measured rate above threshold $i$
* $R_j$ - true (differential) underlying rate in bin $j$ of true $p_T$
* $e_i ^j$ - turn-on efficiency in $i^{th}$ bin of true $p_T$ for threshold $j$

Closure tests with $r$ modeled as $A/e^{\alpha p_T ^{thr}}$ or $B/(p_T ^{thr}-b)^\beta$ and $e_i ^j \sim erf(\gamma(p_{T,j} ^{true}-p_{T,i} ^{thr}))$:

```{r message = F, echo = F}
# With the turn-ons we have to deal with the integrated rates above the thresholds
#  bin everything for simplicity and speed

binLikeliConv <- function(par, turnOn, rate) {
    norm1     <- par[1]
    powerLaw1 <- par[2]
    norm2     <- par[3]
    powerLaw2 <- par[4]
    shift     <- par[5]
    nBins     <- length(binning)
    estRate   <- (rate[1]-rate[nBins]) * (binning[2]-binning[1]) * (
        # function has to be normalized so that when integrated from binning[1] to binning[nBins] it gives unity (or norm1 + norm2)
        norm1/exp(binning*powerLaw1) / ( 1/exp(binning[1]*powerLaw1) - 1/exp(binning[nBins]*powerLaw1) ) * powerLaw1 +
        norm2/(binning+shift)^powerLaw2 / ( 1/(binning[1]+shift)^(powerLaw2-1) - 1/(binning[nBins]+shift)^(powerLaw2-1)) * (powerLaw2-1) )

#    sum( (rate - ((t(turnOn) %*% estRate) + rate[nBins]))^2 ) # MSE in place of log likelihood
#    sum( (rate - ((t(turnOn) %*% estRate) + rate[nBins]))^2 / ( ((t(turnOn) %*% estRate) + rate[nBins]) ) ) # gaus fails for small estRate
    sum( ((t(turnOn) %*% estRate) + rate[nBins]) + rate*(log(rate/((t(turnOn) %*% estRate) + rate[nBins])) - 1) ) # poisson likelihood works for any estimated rate
}

# let's now make a closure test with an infinitely sharp turn-on 

pt3       <- runif(300000,0.00001,1)^(-1/2)
binning   <- seq(1,100,0.1)
rate3Fine <- sapply(binning, function(x) sum(pt3 > x) )
nBins     <- length(binning)

erf <- function(x) 2 * pnorm(x * sqrt(2)) - 1 
idealTurnOn <- matrix(rep(0,nBins*nBins), ncol=nBins, nrow=nBins)
for(bin in 1:nBins) for(thr in 1:nBins) idealTurnOn[thr,bin] = (erf(100*(bin-thr+0.5)) + 1)/2

suppressWarnings( fitIdealPowFine <- nlm(binLikeliConv, c(0,1,1,3,0), turnOn = t(idealTurnOn), rate = rate3Fine, hessian=T) )
# printing fitIdealPowFine seems ok

dfXY <- data.frame( x=runif(1e8,1,1000), y=runif(1e8) )
ptExp <- dfXY[ 1/exp(dfXY$x*2) > dfXY$y,"x"]
rateExpFine <- sapply(binning, function(x) sum(ptExp > x) )

suppressWarnings( fitIdealExpFine <- nlm(binLikeliConv, c(1,1,0,3,0), turnOn = t(idealTurnOn), rate = rateExpFine, hessian=T) )
# printing fitIdealExpFine seems ok too

# coarser binning
binning       <- seq(2,80,1)
rate3Coarse   <- sapply(binning, function(x) sum(pt3 > x) )
rateExpCoarse <- sapply(binning, function(x) sum(ptExp > x) )
nBins         <- length(binning)

idealTurnOn <- matrix(rep(0,nBins*nBins), ncol=nBins, nrow=nBins)
for(bin in 1:nBins) for(thr in 1:nBins) idealTurnOn[thr,bin] = (erf(100*(bin-thr+0.5)) + 1)/2

suppressWarnings( fitIdealPowCoarse <- nlm(binLikeliConv, c(0,1,1,3,0), turnOn = t(idealTurnOn), rate = rate3Coarse, hessian=T) )
# printing fitIdealPowCoarse seems ok

suppressWarnings( fitIdealExpCoarse <- nlm(binLikeliConv, c(1,1,0,3,0), turnOn = t(idealTurnOn), rate = rateExpCoarse, hessian=T) )
# printing fitIdealExpCoarse seems ok

nonIdealTurnOn <- matrix(rep(0,nBins*nBins), ncol=nBins, nrow=nBins)
for(bin in 1:nBins) for(thr in 1:nBins) nonIdealTurnOn[thr,bin] = (erf(0.1*(bin-thr+0.5)) + 1)/2

# non-ideal turn-on
rate3CoarseNI   <- nonIdealTurnOn %*% sapply(1:nBins, function(x) sum(pt3   > binning[x] & ifelse(x<nBins, pt3   < binning[x+1], T)) )
rateExpCoarseNI <- nonIdealTurnOn %*% sapply(1:nBins, function(x) sum(ptExp > binning[x] & ifelse(x<nBins, ptExp < binning[x+1], T)) )

suppressWarnings( fitNonIdealPowCoarse <- nlm(binLikeliConv, c(0,1,1,3,0), turnOn = t(nonIdealTurnOn), rate = rate3CoarseNI, hessian=T) )
# printing fitNonIdealPowCoarse is tentatively ok: 2.26 in place of 3

suppressWarnings( fitNonIdealExpCoarse <- nlm(binLikeliConv, c(1,1,0,3,0), turnOn = t(nonIdealTurnOn), rate = rateExpCoarseNI, hessian=T) )
# printing fitNonIdealExpCoarse is ok: 1.967 in place of 2

slimDF  <- with(d, data.frame(true_pT = muPtGen, response_pT = pt) )
slimDF$trueBin <- sapply(slimDF$true_pT, findBin, binning)
pSpec   <- aggregate(slimDF, by=list(bin=slimDF$trueBin), function(x) x )
norms   <- sapply(pSpec$response_pT,length)
counts  <- t(sapply(pSpec$response_pT,function(x) sapply(binning,function(y) sum(unlist(x)>y) )))
turnOn  <- counts/norms

rate3Real   <- t(turnOn[1:nBins,]) %*% sapply(1:nBins, function(x) sum(pt3   > binning[x] & ifelse(x<nBins, pt3   < binning[x+1], T)) )
rateExpReal <- t(turnOn[1:nBins,]) %*% sapply(1:nBins, function(x) sum(ptExp > binning[x] & ifelse(x<nBins, ptExp < binning[x+1], T)) )

suppressWarnings( fitRealPow <- nlm(binLikeliConv, c(1,1,1,3,0), turnOn = turnOn[1:nBins,], rate = rate3Real, hessian=T) )
# printing fitRealPow is not ok: 4.8 in place of 3

suppressWarnings( fitRealExp <- nlm(binLikeliConv, c(1,1,1,3,0), turnOn = turnOn[1:nBins,], rate = rateExpReal, hessian=T) )
# printing fitRealExp is ok: 1.9 in place of 2

suppressWarnings( fitReal <- nlm(binLikeliConv, c(1,1,1,3,0), turnOn = turnOn[1:nBins,], rate = rateThr, hessian=T) )
# printing fitReal is half-ok 
```

Estimate                    | $\gamma$=100, nBins~1K | $\gamma$=100, nBins~100 | $\gamma$=0.1, nBins~100 | [real turn-on](https://kkotov.github.io/emtfPtRegression/2017.02.10/#5)
----------------------------|------------------------|-------------------------|-------------------------|---------------
$A^{set}/A^{fit}$           | 1/`r signif(fitIdealExpFine$estimate[1],2)` , 0/`r signif(fitIdealPowFine$estimate[1],2)` | 1/`r signif(fitIdealExpCoarse$estimate[1],2)` , 0/`r signif(fitIdealPowCoarse$estimate[1],2)` | 1/`r signif(fitRealExp$estimate[1],2)` , 0/`r signif(fitRealPow$estimate[1],2)`
$\alpha^{set}/\alpha^{fit}$ | 2/`r signif(fitIdealExpFine$estimate[2],2)` , 0/`r signif(fitIdealPowFine$estimate[2],2)` | 2/`r signif(fitIdealExpCoarse$estimate[2],2)` , 0/`r signif(fitIdealPowCoarse$estimate[2],2)` | 2/`r signif(fitRealExp$estimate[1],2)` , 0/`r signif(fitRealPow$estimate[1],2)`
$B^{set}/B^{fit}$           | 0/`r signif(fitIdealExpFine$estimate[3],2)` , 1/`r signif(fitIdealPowFine$estimate[3],2)` | 0/`r signif(fitIdealExpCoarse$estimate[3],2)` , 1/`r signif(fitIdealPowCoarse$estimate[3],2)` | 0/`r signif(fitRealExp$estimate[1],2)` , 1/`r signif(fitRealPow$estimate[1],2)`
$\beta^{set}/\beta^{fit}$   | 0/`r signif(fitIdealExpFine$estimate[4],2)` , 3/`r signif(fitIdealPowFine$estimate[4],2)` | 0/`r signif(fitIdealExpCoarse$estimate[4],2)` , 3/`r signif(fitIdealPowCoarse$estimate[4],2)` | 0/`r signif(fitRealExp$estimate[1],2)` , 3/`r signif(fitRealPow$estimate[1],2)`
$b^{set}/b^{fit}$           | 0/`r signif(fitIdealExpFine$estimate[5],2)` , 0/`r signif(fitIdealPowFine$estimate[5],2)` | 0/`r signif(fitIdealExpCoarse$estimate[5],2)` , 0/`r signif(fitIdealPowCoarse$estimate[5],2)` | 0/`r signif(fitRealExp$estimate[1],2)` , 0/`r signif(fitRealPow$estimate[1],2)`

--- &twocol

## Realistic underlying spectrum

Using parametric form for $R$ and turn-ons $e$ from the [last talk](https://kkotov.github.io/emtfPtRegression/2017.02.10/#5) I assess $R$ from $r$

*** =left

```{r, fig.height=5., fig.width=5., fig.align = 'center', comment = NA, message = F, echo = F}

norm1_r     <- fitReal$estimate[1]
powerLaw1_r <- fitReal$estimate[2]
norm2_r     <- fitReal$estimate[3]
powerLaw2_r <- fitReal$estimate[4]
shift_r     <- fitReal$estimate[5]

realRatePlot <- ggplot( data.frame( pT = seq(2,80,0.1),
                                    rate = ratePDF( seq(2,80,0.1),
                                                    norm1_r,
                                                    powerLaw1_r,
                                                    norm2_r,
                                                    powerLaw2_r,
                                                    shift_r
                                                  )
                                  ),
                        aes(x=pT, y=rate) ) +
        geom_line() # + scale_y_log10() + ylim(, binning[nBins])
#
#        scale_colour_manual(name = "",
#                            values = c("r" = "red", "b"="blue"),
#                            labels = c("r"="EMTF",
#                                       "b"=bquote(.(signif(norm1,2)) ~ "/" ~ e^ ~ {.(signif(powerLaw1,2)) ~ p[T]} ~ "+" ~ .(signif(norm2,2)) ~ "/(" ~ p[T] ~ "+" ~ .(signif(shift,2)) ~ ")"^.(signif(powerLaw2,2)) )
#                                      )
#                           ) +
#
```

*** =right

```{r, fig.height=5., fig.width=5., fig.align = 'center', comment = NA, message = F, echo = F}

```

*** =fullwidth

